---
output:
  word_document: default
  html_document: default
---
# MIS503 – Final ProjectFinal 
## Morris, Jeff
### Zillow Home Value Index Analysis

# Exercise 1: Trends in home values in Wake County, North Carolina
## Wake County Home Sales
```{r}
install.packages("tidyverse")
library(tidyverse)
install.packages("lubridate")
library(lubridate)
install.packages("tidyr")
library(tidyr)
install.packages("ggplot2")
library(ggplot2)

library(readr)
SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
```

```{r}
WakeCountySales <- SingleFamilyResidenceSales
as_tibble(WakeCountySales)

WakeCountySales <- WakeCountySales%>%
   select(RegionName,State,CountyName,Metro,`1996-05`,`1997-05`,`1998-05`,`1999-05`,`2000-05`,`2001-05`,`2002-05`,`2003-05`,`2004-05`,`2005-05`,`2006-05`,`2007-05`,`2008-05`,`2009-05`,`2010-05`,`2011-05`,`2012-05`,`2013-05`,`2014-05`,`2015-05`,`2016-05`,`2017-05`,`2018-05`)%>%

filter(State == "NC",CountyName == "Wake County")%>%

rename(`1996` = `1996-05`,`1997`=`1997-05`,`1998`=`1998-05`,`1999` =`1999-05`,`2000`= `2000-05`,`2001`= `2001-05`,`2002`=`2002-05`,`2003`=`2003-05`,`2004`=`2004-05`,`2005`=`2005-05`,`2006`=`2006-05`,`2007`=`2007-05`,`2008`=`2008-05`,`2009`=`2009-05`,`2010`=`2010-05`,`2011`=`2011-05`,`2012`=`2012-05`,`2013`=`2013-05`,`2014`=`2014-05`,`2015`=`2015-05`,`2016`=`2016-05`,`2017`=`2017-05`,`2018`=`2018-05`)%>%

gather(key= YR,value = ZHVI, "1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013", "2014","2015","2016","2017","2018", na.rm = FALSE,convert = FALSE)

WakeCountySales

```

```{r}
ggplot(WakeCountySales, aes(x=YR, y=ZHVI, color=RegionName)) +
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)+
  ggtitle("Wake County Home Sales") +
  xlab ("Year") + ylab ("Home Values")
  
```

```{r}
ggplot(WakeCountySales, aes(x=YR, y=ZHVI, color=RegionName)) +
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)+
  ggtitle("Wake County Home Sales") +
  xlab ("Year") + ylab ("Home Values")
  
```




### Question a. What have been the overall trends in Wake County Home Values?
### Answer a. Three peaks and two valleys, with an overall trend of increased value 

### Question b. There were dips in home values in the past 20 years. What years did these occur?
### Answer b. Between 1998-1999 and 2011-2012

### Question c. Based on the analysis, where would be the least expensive area to purchase home? Most expensive area?
### Answer c. 2018 least expensive area to buy home is Zebulon and most expensive area to buy a home is Raleigh.  

### Question d. Are any area home values trending down? Is there one area that stands out compared to others?
### Answer d. Over the last 22 years all home values are trending up, specifically in Raleigh there appears to be spike in 2018 in home sale values 

# Exercise 2: North Carolina Rental Market
## NC Rental Market 

```{r}
install.packages("tidyverse")
library(tidyverse)
install.packages("lubridate")
library(lubridate)
install.packages("tidyr")
library(tidyr)
install.packages("ggplot2")
library(ggplot2)

library(readr)
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
```

```{r}
Rentals <- SingleFamilyResidenceRental
as_tibble(Rentals)

Rentals <- Rentals%>%
   select(RegionName,State,`2010-11`,`2011-11`,`2012-11`,`2013-11`,`2014-11`,`2015-11`,`2016-11`,`2017-11`,`2018-10`)%>%
  filter(State == 'NC', RegionName %in% c('Asheville','Charlotte','Durham','Fayetteville','Raleigh','Wilmington'))%>%
  rename(`2010`=`2010-11`,`2011`=`2011-11`,`2012`=`2012-11`,`2013`=`2013-11`,`2014`=`2014-11`,`2015`=`2015-11`,`2016`=`2016-11`,`2017`=`2017-11`,`2018`=`2018-10`)%>%

  gather(key= YR,value = ZHVI,"2010","2011","2012","2013", "2014","2015","2016","2017","2018", na.rm = FALSE,convert = FALSE)

Rentals
```

```{r}
ggplot(Rentals, aes(x=YR, y=ZHVI, color=RegionName)) +
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Rental Values (in $)", labels = scales::comma)+
  ggtitle("Rental Costs in NC Cities") +
  xlab ("Year") + ylab ("Average Rental Cost")
  
```

```{r}
ggplot(Rentals, aes(x=YR, y=ZHVI, color=RegionName)) +
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Rental Values (in $)", labels = scales::comma)+
  ggtitle("Rental Costs in NC Cities") +
  xlab ("Year") + ylab ("Average Rental Cost")
```




### Question a. What has been the overall trend in the rental market around the state? Are there any cities that have not followed this trend?
### Answer a. Upward trend for overall rental values; however, Fayetteville is the exception showing a downward trend.  



### Question b. Where is the most expensive city to rent in? Least expensive?
### Answer b. Most expensive to rent is Asheville, least expensive is Fayetteville.  


### Question c. You are trying decide between Wilmington and Asheville. Which market has the lowest rent?
### Answer c. Wilmington has a lower rent than Asheville.  

# Exercise 3: Home Values in Select North Carolina Markets
## Home Values in Select Rental Markets 

```{r}
NCHomeSales <- SingleFamilyResidenceSales
as_tibble(NCHomeSales)

NCHomeSales <- NCHomeSales%>%
   select(RegionName,Metro,State,`1996-05`,`1997-05`,`1998-05`,`1999-05`,`2000-05`,`2001-05`,`2002-05`,`2003-05`,`2004-05`,`2005-05`,`2006-05`,`2007-05`,`2008-05`,`2009-05`,`2010-05`,`2011-05`,`2012-05`,`2013-05`,`2014-05`,`2015-05`,`2016-05`,`2017-05`,`2018-05`)%>%
  filter(State == 'NC', Metro %in% c('Asheville','Charlotte-Concord-Gastonia','Raleigh','Wilmington'))%>%
  rename(`1996` = `1996-05`,`1997`=`1997-05`,`1998`=`1998-05`,`1999` =`1999-05`,`2000`= `2000-05`,`2001`= `2001-05`,`2002`=`2002-05`,`2003`=`2003-05`,`2004`=`2004-05`,`2005`=`2005-05`,`2006`=`2006-05`,`2007`=`2007-05`,`2008`=`2008-05`,`2009`=`2009-05`,`2010`=`2010-05`,`2011`=`2011-05`,`2012`=`2012-05`,`2013`=`2013-05`,`2014`=`2014-05`,`2015`=`2015-05`,`2016`=`2016-05`,`2017`=`2017-05`,`2018`=`2018-05`)%>%
  gather(key= YR,value = ZHVI, "1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013", "2014","2015","2016","2017","2018", na.rm = FALSE,convert = FALSE)

NCHomeSales
```

```{r}
ggplot(NCHomeSales, aes(x=Metro, y=ZHVI)) +
  geom_violin()+ geom_boxplot(width=0.2)+
  ggtitle("Home Values Areas of Interest") +
  xlab ("Metro") + ylab ("ZHVI")
```

### Question a. According to the results, which market has the lowest median price (represented as horizontal bar in box plot)?
### Answer b. Charlotte-Concord-Gastonia

### Queston b. The violin plot will show density meaning the wider the plot is, the more observations occur within that area. Which market has the most density around the median value of homes?
### Answer b. Charlotte-Concord-Gastonia

# Exercise 4: Relocation to Another City
## Relocation Home Value Comparison

```{r}
NationalHomeSales <- SingleFamilyResidenceSales
as_tibble(NationalHomeSales)

NationalHomeSales <- NationalHomeSales%>%
   select(RegionName,State,Metro,`1996-05`,`1997-05`,`1998-05`,`1999-05`,`2000-05`,`2001-05`,`2002-05`,`2003-05`,`2004-05`,`2005-05`,`2006-05`,`2007-05`,`2008-05`,`2009-05`,`2010-05`,`2011-05`,`2012-05`,`2013-05`,`2014-05`,`2015-05`,`2016-05`,`2017-05`,`2018-05`)%>%
  filter(State%in% c('IL',"CO", "TX", "NY"),RegionName%in% c('Chicago','Denver','Houston','New York'))%>%
  rename(`1996` = `1996-05`,`1997`=`1997-05`,`1998`=`1998-05`,`1999` =`1999-05`,`2000`= `2000-05`,`2001`= `2001-05`,`2002`=`2002-05`,`2003`=`2003-05`,`2004`=`2004-05`,`2005`=`2005-05`,`2006`=`2006-05`,`2007`=`2007-05`,`2008`=`2008-05`,`2009`=`2009-05`,`2010`=`2010-05`,`2011`=`2011-05`,`2012`=`2012-05`,`2013`=`2013-05`,`2014`=`2014-05`,`2015`=`2015-05`,`2016`=`2016-05`,`2017`=`2017-05`,`2018`=`2018-05`)%>%
  gather(key= YR,value = ZHVI, "1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013", "2014","2015","2016","2017","2018", na.rm = FALSE,convert = FALSE)

NationalHomeSales
```

```{r}
ggplot(NationalHomeSales, aes(x=YR, y=ZHVI)) +
  geom_point()+
  facet_wrap(~Metro)+
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size=5))+
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)+
  ggtitle("City Home Value Comparison") +
  xlab ("Year") + ylab ("Home Values")
```

### Question a. Based on your analysis, which city’s housing is most affordable? Least affordable?
### Answer a.  Most affordable Houston-The Woodlands-Sugar Land, Least affordable New York-Newark-Jersey City 

### Question b. Which cities saw the largest change in prices over the past 5 years? Which city has remained more consistent (i.e., no huge swings up or down in home values)?
### Answer b. Largest change occurred in Denver and New York.  Houston had the most consistent prices.  

### Question c. During the market downturn in 2012, which cities were most impacted? Which cities have recovered?
### Answer c. Chicago and New York were most impacted, New York has recovered, Chicago has still not returned to housing prices prior to 2012




